---
title: "Take-Home Exercise 3"
description: |

  Putting Visual Analytics into Practical Use

author:
  - name: Lee Xiao Qi 
    url: https://example.com/norajones
    affiliation: School of Computing and Information Systems (SMU)
    affiliation_url: https://example.com/spacelysprokets
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

# 1. The Task

We are required to attempt one of the three questions under Challenge 3 of [VAST Challenge 2022](https://vast-challenge.github.io/2022/) on the economic of the city of Engagement, Ohio USA by using appropriate static and interactive statistical graphics methods.

With that, I have chosen #3 which is to:

**Describe the health of the various employers within the city limits. What employment patterns do you observe? Do you notice any areas of particularly high or low turnover? Limit your response to 10 images and 500 words.**

# 2. The Datasets

### 2.1 About the Data

About 1000 representative residents have volunteered to provide data using the city’s urban planning app, which records the places they visit, their spending, and their purchases, among other things; totaling to 3 dataset folders (**Activity Logs, Attributes, Journals**). For this exercise, we refers to Empolyers that offers jobs as stipulated in the **Jobs** dataset; we thus selected the dataset **Employers**, **Buildings**, **Jobs** & **Participants** from the **Attributes** folder, and **CheckinJournal** from the **Journals** folder which give details on the employers, employees and jobs.  

### 2.2 Importing the relevant packages and data

The packages **tidyverse** (including **dplyr**, **ggplot2**, **patchwork**), **ggrepel**, **lubridate**, **gapminder**, **gganimate**, **ggiraph** etc will be used for the purpose of this exercise:

The code chunk below is used to install and load the required packages onto RStudio.

```{r}
packages = c('tidyverse','treemap','ggrepel','lubridate','gapminder','gganimate','ggiraph','plotly','zoo')

for(p in packages){
  if(!require(p, character.only =T)){
    install.packages(p)
    }
  library(p, character.only =T)
}
```

The code chuck below import *Employers.csv*, *Buildings.csv*, *Jobs.csv*, *CheckinJournal.csv* and *Participants.csv* from the data folder into R by using `read_csv()` and save it as an tibble data frame.  

```{r}
Employers <- read_csv("data/Employers.csv")
Buildings <- read_csv("data/Buildings.csv")
Jobs <- read_csv("data/Jobs.csv")
Checkin <- read_csv("data/CheckinJournal.csv")
Participants <- read_csv("data/Participants.csv")

```
### 2.3 Data Exploration and Wrangling

First, let’s get a general sense of our data using the function `summary()`.

```{r}
summary(Employers)
summary(Buildings)
summary(Jobs)
summary(Checkin)

```
#### 2.3.1 Joining of Datasets

*Employers* dataset is joined with *Buildings* dataset based on `buildingId` to filter only relevant information from *Buildings* dataset in regards to employers. Left join on *Employers* is used as commercial buildings are a subset of the different type of buildings.  

```{r}
Employers <- Employers %>% left_join(Buildings,by="buildingId")

```

Next, we use outer join on *Employers* and *Jobs* based on `employerId` to have a full overview on all the jobs that are offered by each employer.   

```{r}
Employers = merge(x=Employers,y=Jobs,by="employerId",all=TRUE)

```

#### 2.3.2 Renaming of Columns and Values

First, we start by renaming the columns and values of in *Employers* and *Checkin* dataset using the function [`rename()`](https://dplyr.tidyverse.org/reference/rename.html), and [`sub()`](https://www.datasciencemadesimple.com/sub-gsub-function-in-r/) for a better format and ease of reading.

Note: A check between both datasets shows that `venueId` in *Checkin* dataset refers to the `employerId`, `pubId` etc. For the purpose of this exercise, we are only interested in the `employerId` (`venueType` = `Workspace`) and other venues type will be removed subsequently.

```{r}
# rename columns
Employers <- Employers %>%
  rename('Employer_ID' = 'employerId', 
         'Location(Pt)' = 'location.x', 
         'Location(Area)' = 'location.y',
         'Building_ID' = 'buildingId', 
         'Building_Type' = 'buildingType', 
         'Max_Occupancy' = 'maxOccupancy', 
         'Units' = 'units', 
         'Job_ID' = 'jobId',
         'Hourly_Rate' = 'hourlyRate',
         'Start_Time' = 'startTime',
         'End_Time' = 'endTime',
         'Days_To_Work' = 'daysToWork',
         'Education_Level' = 'educationRequirement')

Checkin <- Checkin %>%
  rename('Participant_ID' = 'participantId', 
         'Timestamp' = 'timestamp', 
         'Employer_ID' = 'venueId',
         'Venue_Type' = 'venueType')

Jobs <- Jobs %>%
  rename('Education_Level' = 'educationRequirement')


Participants <- Participants %>%
  rename('Participant_ID' = 'participantId', 
         'Household_Size' = 'householdSize', 
         'Have_Kids' = 'haveKids', 
         'Age' = 'age', 
         'Education_Level' = 'educationLevel', 
         'Interest_Group' = 'interestGroup', 
         'Joviality' = 'joviality')

#rename row values
Employers$Education_Level <- sub('HighSchoolOrCollege', 
                                    'High School or College',
                                    Employers$Education_Level)
Jobs$Education_Level <- sub('HighSchoolOrCollege', 
                                    'High School or College',
                                    Jobs$Education_Level)
Participants$Education_Level <- sub('HighSchoolOrCollege', 
                                    'High School or College',
                                    Participants$Education_Level)

```

#### 2.3.3 Fliter Data, Compute Percentage Change in Number of Employee Over Time 

We are using *Checkin* to see the changes in employment (i.e. checkin by participants at workplace) over time. We see that the `venueId` column in the dataset are IDs of all possible venues such as work place, restaurants and pubs. 

Given that we are only interested in workplace, we will first filter out row that reads "workplace" using `grep()`. 

We then assign a running week number and also day to day percentage change in number of employees for each employer. 

```{r}
#Extract the date from timestamp
Checkin$Date <- as.Date(Checkin$Timestamp)

#Filter rows with workplace as value
Workplace_Checkin <- Checkin[grep("Workplace", Checkin$Venue_Type),]

#Assign Running Week Number
Workplace_Checkin <- Workplace_Checkin %>%
  mutate(Week_Num = as.double(ceiling(difftime(Workplace_Checkin$Date, strptime("01.03.2022", format = "%d.%m.%Y"),units="weeks"))))

#Compute no. of employees that report to work during that week (5day work week) 
Count_Checkin <- Workplace_Checkin %>%
  group_by(Week_Num, Employer_ID) %>%
  summarise('Num_of_Employees'= n_distinct(Participant_ID)) %>%  
  ungroup()

#Calculate Percentage Change  
Count_Checkin <- Count_Checkin %>%  
  group_by(Employer_ID) %>%
  mutate(
    Perc_Chg = round((Num_of_Employees - lag(Num_of_Employees))/lag(Num_of_Employees)*100,2)
  )

```
# 3. Visualizing with Charts

Next, we put up a series of charts to address the questions.

### 3.1 Turnover Rate

We see the absolute and percentage change in number of employee for each employer over each work week 

```{r}
Count_Checkin$Employer_ID <- as.character(Count_Checkin$Employer_ID)


p<- ggplot(Count_Checkin, aes(x=Week_Num, y=Num_of_Employees, group=Employer_ID)) +
  geom_line(aes(color=Employer_ID),show.legend = FALSE)+
  ylim(1,15) +
  ggtitle(label = "Number of Employees For Each Work Week")+
  xlab("Week Number from Mar 22 to May 23") +
  ylab("No. of Employees") +
  theme_minimal()+
  theme(plot.title = element_text(size=12, face="bold",hjust = 0.5))  

ggplotly(p)
```
Most of the manpower movement happens on the first two weeks where alot employement dropped, noting sharp decline for Employer 1763 and 868; which started off with 14 and 15 in Week 1 and then dropped to 6 employees in Week 2.

```{r}

p<- ggplot(Count_Checkin, aes(x=Week_Num, y=Perc_Chg,group=Employer_ID)) +
  geom_line(aes(color=Employer_ID), show.legend = FALSE)+
  ggtitle(label = "Percentage % Change of Employees For Each Work Week")+
  xlab("Week Number from Mar 22 to May 23") +
  ylab("% Change") +
  theme_minimal()+
  theme(plot.title = element_text(size=12, face="bold",hjust = 0.5)) 


ggplotly(p)
```
Similarly, most of the decline happened in the first two weeks and remained stabled mostly throughout untill the last week.  

```{r,echo=FALSE}
library(magrittr)
```

### 3.2 Employment Pattern - Wage vs Education vs Age Group

Next, we want to know the age profile of employees for each employer. We grouped them into various age groups for easier readability. 

We use left join to find out the employer for participants and the respective age group they are in.

```{r}
Participants_AgeGrp <- Participants %>%
   select(Participant_ID, Age, Education_Level)

Workplace_Checkin<-
  left_join(Workplace_Checkin,Participants_AgeGrp,by = 'Participant_ID') 


```

We now use raincloud plot to see the wage distribution for different education level.

```{r,fig.width=14,fig.height=12}

Jobs$Education_Level = factor(Jobs$Education_Level, levels = c('Low', 'High School or College', 'Bachelors','Graduate'))

p<- ggplot(Jobs, aes(x = Education_Level, y = hourlyRate, fill=Education_Level)) + 
  ggdist::stat_halfeye(
    adjust = .5, 
    width = .6, 
    .width = 0, 
    justification = -.3, 
    point_colour = NA) + 
  geom_boxplot(
    width = .25, 
    outlier.shape = NA
  ) +
  geom_point(
    size = 1.3,
    alpha = .3,
    position = position_jitter(
      seed = 1, width = .1
    )
  ) + 
  coord_cartesian(xlim = c(1.2, NA), clip = "off")+
  ggtitle(label = "Wage Distribution for Different Education Level",
          subtitle = "High Wages For Higher Educated")+
  theme_minimal()+
  theme(plot.title = element_text(size=14, face="bold",hjust = 0.5),
          plot.subtitle = element_text(size=12,hjust = 0.5,color='mediumvioletred'))+
  theme(axis.title.y= element_text(angle=0), axis.ticks.x= element_blank(),
        panel.background= element_blank(), axis.line= element_line(color= 'grey'))

ggplotly(p)
```
We see that indeed higher education are paid better. 

We use a heatmap with `treepmap` to see the age profile for each employer. 

```{r,echo=FALSE}
library(d3treeR)
```

```{r,fig.width=20, fig.height=5, echo=FALSE}

tm <- treemap(Workplace_Checkin,
        index=c("Employer_ID"),
        vSize="Participant_ID",
        vColor="Age",
        type = "dens")

```
```{r,fig.width=20, fig.height=5}

d3tree(tm, rootname = "Age and No. of Employees under Each Employer" )
```